---
title: "Strng and factors"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

library(p8105.datasets)
library(rvest)


knitr::opts_chunk$set(
	fig.asp = 0.6,
	fig.width = 6,
	out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continous.color = "viridis",
  ggplot2.continous.fill = "viridis"
)

scale_color_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Strings and regex

```{r}

string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

```

For exact matches, you can designate matches at the beginning or end of a line

```{r}

string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think") #^ denotes that you want to search for something that starts with i think
str_detect(string_vec, "i think$") #$denotes that you want to search for something that ends with i think

```

```{r}

string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")

```

```{r}

string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")

```

The character . matches anything.

```{r}

string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # need to put \\ to search character to find something with a . "7\\.11"

```

```{r}

string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")

```

## Factors

```{r}

factor_vec = factor(c("male","male","female","female"))

factor_vec

as.numeric(factor_vec)



```

what happens if i relevel....


```{r}

factor_vec = fct_relevel(factor_vec, "male")

as.numeric(factor_vec)

```



## NSDUH-- strings

```{r}

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"


table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()

```


```{r}

data_marj =
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State, #Chooses everything except state column
    names_to = "age_year",
    values_to = "percent"
  ) %>%
  separate(age_year, into = c("age","year"), sep = "\\(", ) %>%
  mutate(
    year = str_replace(year,"\\)",""),
    percent = str_replace(percent, "[a-c]$",""), #specifying what to remove
    percent = as.numeric(percent)
  ) %>%
  filter(!(State %in% c("Total U.S.","Northeast","Midwest","South","West")))

```


## NSDUH -- factors


```{r}

data_marj %>%
  filter(age == "12-17") %>%
  mutate(State = fct_reorder(State,percent)) %>% #change state to factor, then reorder by percent
  ggplot(aes(State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


## Weather data

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```


```{r}

weather_df %>%
  mutate(name = fct_reorder(name, tmax)) %>%
  ggplot(aes(x= name, y = tmax)) +
  geom_violin()

```


```{r}

weather_df %>%
  lm(tmax ~ name, data = .)

weather_df %>%
  mutate(name = fct_reorder(name, tmax)) %>%
  lm(tmax ~ name, data = .)

```





## Resturant Inspections

```{r}

data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

```

To simplify, remove inspection scores other than A, B, C and remove the resturants with missing boro info

```{r}

rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))

```

focus on pizza places for now and re-examine grades by borough

```{r}

rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

```

Need to tell it not to be case sensitive

```{r}

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

```

Visualize

```{r}

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

```

Try a different order- number of pizza places using fct_infreq

```{r}

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 


```

Can rename borough using str_replace

```{r}

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Brooklyn", "HipsterHQ")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 

```


Can try to use base R replace


```{r}

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Brooklyn"), "HipsterHQ")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

```

Didn't work- can use dedicated function to rename factor levels


```{r}

rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "HipsterHQ" = "Brooklyn")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

```










